---
title: "BDA - Project"
author: "Huy Nguyen, Matias Nieminen, Zehang Li"
output:
  html_document:
    toc: yes
    toc_depth: '1'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 1
urlcolor: blue
---

```{r, message=FALSE}
library(aaltobda)
library(bayesplot)
library(cmdstanr)
library(plyr)
library(dplyr)
library(ggplot2)
library(ggdist) # for stat_dotsinterval
library(posterior)
library(brms)
# Set more readable themes with bigger font for plotting packages.
ggplot2::theme_set(theme_minimal(base_size = 14))
bayesplot::bayesplot_theme_set(theme_minimal(base_size = 14))
# This registers CmdStan as the backend for compiling cmdstan-chunks.
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)
register_knitr_engine(override = FALSE)
```

The data comes from `dslabs` library. In the preprocessing step, datapoints containing NA values are removed. The number of people voting for Clinton and Trump (integers) are then calculated.
```{r}
library(dslabs)
data("polls_us_election_2016")

rawdata = na.omit(subset(polls_us_election_2016, 
                         select=c(state, samplesize, rawpoll_clinton, rawpoll_trump)))

y_clinton = as.integer(rawdata$samplesize*rawdata$rawpoll_clinton/100)
y_trump = as.integer(rawdata$samplesize*rawdata$rawpoll_trump/100)

election_data =  data.frame(as.integer(rawdata$state), rawdata$samplesize, y_clinton, y_trump)
names(election_data) = c("state","samplesize","y_clinton","y_trump")
```

Here, the 3 CmdStan models are trained, and the MCMC Inference is done with the `⁠$sample()`⁠ method.
```{r, message=FALSE}
stan_data = list(N_observations=nrow(election_data),
                 N_states=length(unique(election_data$state)), 
                 samplesize = election_data$samplesize,
                 state_idx = election_data$state, 
                 y_clinton = election_data$y_clinton,
                 y_trump = election_data$y_trump)

pooled_model = cmdstan_model("project_pooled_model.stan")
fit_pooled = pooled_model$sample(data = stan_data, refresh=0, show_messages=FALSE)

separate_model = cmdstan_model("project_separate_model.stan")
fit_separate = separate_model$sample(data = stan_data, refresh=0, show_messages=FALSE)

hierarchical_model = cmdstan_model("project_hierarchical_model.stan")
fit_hierarchical = hierarchical_model$sample(data = stan_data, refresh=0, show_messages=FALSE)
```

Draws from the 3 models for `theta_clinton[2]`:
```{r, message=FALSE}
ndraws <- nrow(fit_pooled$sampler_diagnostics(format = "matrix"))
posterior_theta_clinton_1 <- data.frame(
  model_name = rep(c("Pooled", "Separate", "Hierarchical"), each = ndraws),
  theta_clinton_1 = c(extract_variable(fit_pooled, "theta_clinton[1]"), 
                      extract_variable(fit_separate, "theta_clinton[1]"), 
                      extract_variable(fit_hierarchical, "theta_clinton[1]")))

predictive_theta_clinton_1 <- data.frame(
  model_name = rep(c("Pooled", "Separate", "Hierarchical"), each = ndraws),
  theta_clinton_pred_1 = c(extract_variable(fit_pooled, "theta_clinton_pred[1]"),
                           extract_variable(fit_separate, "theta_clinton_pred[1]"), 
                           extract_variable(fit_hierarchical, "theta_clinton_pred[1]")))
```

And plot them:
```{r, fig.height=5, fig.width=8}
ggplot(posterior_theta_clinton_1, aes(x = theta_clinton_1, y = model_name)) +
  stat_dotsinterval(quantiles = 100, scale = .9) +
  labs(title = "posterior", x = "theta_clinton_1", y = "Model")
```

```{r, fig.height=5, fig.width=8}
ggplot(predictive_theta_clinton_2, aes(x = theta_clinton_pred_2, y = model_name)) +
  stat_dotsinterval(quantiles = 100, scale = .9) +
  labs(title = "posterior predictive", x = "theta_clinton_pred_2", y = "Model")
```

Draws from the 3 models for all `theta_clinton`:
```{r}
N_states=length(unique(election_data$state))
theta_clinton_pooled_list = data.frame(number=1:4000)
theta_clinton_separate_list = data.frame(number=1:4000)
theta_clinton_hierarchical_list = data.frame(number=1:4000)
for (i in 1:N_states){
  variable = paste("theta_clinton[",toString(i),"]",sep="")
  theta_clinton_pooled = extract_variable(fit_pooled, variable)
  theta_clinton_separate = extract_variable(fit_separate, variable)
  theta_clinton_hierarchical = extract_variable(fit_hierarchical, variable)
  
  theta_clinton_pooled_list[variable] = theta_clinton_pooled
  theta_clinton_separate_list[variable] = theta_clinton_separate
  theta_clinton_hierarchical_list[variable] = theta_clinton_hierarchical
}
theta_clinton_pooled_list = subset(theta_clinton_pooled_list, select=-c(number))
theta_clinton_separate_list = subset(theta_clinton_separate_list, select=-c(number))
theta_clinton_hierarchical_list = subset(theta_clinton_hierarchical_list, select=-c(number))
```

And `theta_clinton_pred`:
```{r}
N_states=length(unique(election_data$state))
theta_clinton_pred_pooled_list = data.frame(number=1:4000)
theta_clinton_pred_separate_list = data.frame(number=1:4000)
theta_clinton_pred_hierarchical_list = data.frame(number=1:4000)
for (i in 1:N_states){
  variable = paste("theta_clinton_pred[",toString(i),"]",sep="")
  theta_clinton_pred_pooled = extract_variable(fit_pooled, variable)
  theta_clinton_pred_separate = extract_variable(fit_separate, variable)
  theta_clinton_pred_hierarchical = extract_variable(fit_hierarchical, variable)
  
  theta_clinton_pred_pooled_list[variable] = theta_clinton_pred_pooled
  theta_clinton_pred_separate_list[variable] = theta_clinton_pred_separate
  theta_clinton_pred_hierarchical_list[variable] = theta_clinton_pred_hierarchical
}
theta_clinton_pred_pooled_list = subset(theta_clinton_pred_pooled_list, select=-c(number))
theta_clinton_pred_separate_list = subset(theta_clinton_pred_separate_list, select=-c(number))
theta_clinton_pred_hierarchical_list = subset(theta_clinton_pred_hierarchical_list, select=-c(number))
```

And plot them:
```{r, fig.height=6, fig.width=12}
plot(1:N_states, colMeans(theta_clinton_pooled_list), 
     xlab="state id", ylab="theta_clinton", col="red", ylim=c(0.2,0.8), type="l")
lines(1:N_states, colMeans(theta_clinton_separate_list), col="blue", type="l")
lines(1:N_states, colMeans(theta_clinton_hierarchical_list), col="orange", type="l")

lines(1:N_states, colMeans(theta_clinton_pred_pooled_list), col="red", type="l", lty=2)
lines(1:N_states, colMeans(theta_clinton_pred_separate_list), col="blue", type="l", lty=2)
lines(1:N_states, colMeans(theta_clinton_pred_hierarchical_list), col="orange", type="l", lty=2)

legend(45, 0.8, legend=c("posterior_pooled", "posterior_separate", "posterior_hierarchical", "predictive_pooled", "predictive_separate", "predictive_hierarchical"),
       col=c("red","blue","orange", "red","blue","orange"), lty=c(1,1,1,2,2,2))
```

# Convergence analysis with Rhat and ESS
```{r}
summary_pooled = fit_pooled$summary()
rhat_values_pooled = summary_pooled[, "rhat"]
ess_pooled = summary_pooled[, "ess_tail"]
print(max(rhat_values_pooled))
print(min(ess_pooled, na.rm=TRUE))
```

```{r}
summary_separate = fit_separate$summary()
rhat_values_separate = summary_separate[, "rhat"]
ess_separate = summary_separate[, "ess_tail"]
print(max(rhat_values_separate))
print(min(ess_separate, na.rm=TRUE))
```

```{r}
summary_hierarchical = fit_hierarchical$summary()
rhat_values_hierarchical = summary_hierarchical[, "rhat"]
ess_hierarchical = summary_hierarchical[, "ess_tail"]
print(max(rhat_values_hierarchical))
print(min(ess_hierarchical, na.rm=TRUE))
```

# Model comparison with LOO-CV
```{r}
loo_pooled = fit_pooled$loo(variables = "theta_clinton")
loo_pooled
```

```{r}
loo_separate = fit_separate$loo(variables = "theta_clinton")
loo_separate
```

```{r}
loo_hierarchical = fit_hierarchical$loo(variables = "theta_clinton")
loo_hierarchical
```

```{r}
loo_compare(loo_pooled, loo_separate, loo_hierarchical)
```

# Sensitivity analysis
Pooled and separate: Beta(9,1) and Beta(1,9) instead of Beta(2,2) and Beta(9,1)
Hierarchical: Uniform(10,20) instead of Uniform(0,10)
```{r, message=FALSE}
pooled_model_2 = cmdstan_model("project_pooled_model - Copy.stan")
fit_pooled_2 = pooled_model_2$sample(data = stan_data, refresh=0, show_messages=FALSE)

separate_model_2 = cmdstan_model("project_separate_model - Copy.stan")
fit_separate_2 = separate_model_2$sample(data = stan_data, refresh=0, show_messages=FALSE)

hierarchical_model_2 = cmdstan_model("project_hierarchical_model - Copy.stan")
fit_hierarchical_2 = hierarchical_model_2$sample(data = stan_data, refresh=0, show_messages=FALSE)
```

```{r, message=FALSE}
theta_clinton_2_2 = c(extract_variable(fit_pooled_2, "theta_clinton[2]"), 
                     extract_variable(fit_separate_2, "theta_clinton[2]"), 
                     extract_variable(fit_hierarchical_2, "theta_clinton[2]"))
posterior_theta_clinton_2_2 <- data.frame(
  model_name = rep(c("Pooled_2", "Separate_2", "Hierarchical_2"), each = ndraws),
  theta_clinton_2_2)

theta_clinton_pred_2_2 = c(extract_variable(fit_pooled_2, "theta_clinton_pred[2]"),
                            extract_variable(fit_separate_2, "theta_clinton_pred[2]"), 
                            extract_variable(fit_hierarchical_2, "theta_clinton_pred[2]"))
predictive_theta_clinton_2_2 <- data.frame(
  model_name = rep(c("Pooled_2", "Separate_2", "Hierarchical_2"), each = ndraws),
  theta_clinton_pred_2_2)
```

And plot them:
```{r, fig.height=5, fig.width=8}
ggplot(posterior_theta_clinton_2_2, aes(x = theta_clinton_2_2, y = model_name)) +
  stat_dotsinterval(quantiles = 100, scale = .9) +
  labs(title = "posterior - sensitivity analysis", x = "theta_clinton_2", y = "Model")
```
```{r, fig.height=5, fig.width=8}
ggplot(predictive_theta_clinton_2_2, aes(x = theta_clinton_pred_2_2, y = model_name)) +
  stat_dotsinterval(quantiles = 100, scale = .9) +
  labs(title = "posterior predictive - sensitivity analysis", x = "theta_clinton_pred_2", y = "Model")
```


















